---
title: "R Notebook"
output: html_notebook
---

# Stats 2 Project 

Members: Daniel Chang, Carolina Craus, Andrew Yule

## Data 

[Bank Marketing Data Set:](https://archive.ics.uci.edu/ml/datasets/Bank+Marketing) 
The data is related with direct marketing campaigns (phone calls) of a Portuguese banking institution. The classification goal is to predict if the client will subscribe (yes/no) to a term deposit (variable y). 

Input variables:
# bank client data:
1 - age (numeric)
2 - job : type of job (categorical: 'admin.','blue-collar','entrepreneur','housemaid','management','retired','self-employed','services','student','technician','unemployed','unknown')
3 - marital : marital status (categorical: 'divorced','married','single','unknown'; note: 'divorced' means divorced or widowed)
4 - education (categorical: 'basic.4y','basic.6y','basic.9y','high.school','illiterate','professional.course','university.degree','unknown')
5 - default: has credit in default? (categorical: 'no','yes','unknown')
6 - housing: has housing loan? (categorical: 'no','yes','unknown')
7 - loan: has personal loan? (categorical: 'no','yes','unknown')
# related with the last contact of the current campaign:
8 - contact: contact communication type (categorical: 'cellular','telephone')
9 - month: last contact month of year (categorical: 'jan', 'feb', 'mar', ..., 'nov', 'dec')
10 - day_of_week: last contact day of the week (categorical: 'mon','tue','wed','thu','fri')
11 - duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.
# other attributes:
12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
14 - previous: number of contacts performed before this campaign and for this client (numeric)
15 - poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')
# social and economic context attributes
16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
17 - cons.price.idx: consumer price index - monthly indicator (numeric)
18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric)
19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
20 - nr.employed: number of employees - quarterly indicator (numeric)

Output variable (desired target):
21 - y - has the client subscribed a term deposit? (binary: 'yes','no')

## Objectives 

1. Perform EDA and build a logistic regression model for interpretation purposes. 

2. With a simple logistic regression model as a baseline, perform additional competing models to improve on prediction perfromance metrics. 

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Data Exploration 

Load in Data
```{r}
bank <- read.csv("C:/Users/caroc/OneDrive/Desktop/Stat2-Project2/bank-full.csv", stringsAsFactors=TRUE)
head(bank)
```

```{r}
summary(bank)
```

```{r}
str(bank)
```
Check for missing values 
No NAs in data
```{r}
colSums(is.na(bank))
```
```{r}
bank %>% 
  summarise_all(list(~sum(. == "unknown"))) %>% 
  gather(key = "variable", value = "nr_unknown") %>% 
  arrange(-nr_unknown)
```
## EDA 
```{r}
# Contact 
ggplot(data = bank, aes(x = contact , y = y, fill = y)) + 
  geom_bar(stat = 'identity', position = 'dodge') + xlab("Contact")+ylab("Response")
# Education 
ggplot(data = bank, aes(x = education , y = y, fill = y)) + 
  geom_bar(stat = 'identity', position = 'dodge') + xlab("Education")+ylab("Response")
# Job 
ggplot(data = bank, aes(x = y, y = job, fill = y)) + 
  geom_bar(stat = 'identity', position = 'dodge') + xlab("Job")+ylab("Response") +
  theme(axis.title.x = element_text(size = 6))

# Marital Status 
ggplot(data = bank, aes(x = marital, y = y, fill = y)) + 
  geom_bar(stat = 'identity', position = 'dodge') + xlab("Marital")+ylab("Response")

# Default
ggplot(data = bank, aes(x = default , y = y, fill = y)) + 
  geom_bar(stat = 'identity', position = 'dodge') + xlab("Default")+ylab("Response")

#Balance 
ggplot(data = bank, aes(x = balance , y = y, fill = y)) + 
  geom_bar(stat = 'identity', position = 'dodge') + xlab("Balance")+ylab("Response")

# Housing 
ggplot(data = bank, aes(x = housing , y = y, fill = y)) + 
  geom_bar(stat = 'identity', position = 'dodge') + xlab("Housing")+ylab("Response")
# Loan 
ggplot(data = bank, aes(x = loan , y = y, fill = y)) + 
  geom_bar(stat = 'identity', position = 'dodge') + xlab("Loan")+ylab("Response")
```

